---
title: "ES 193DS homework 3"
date: 2024-06-02
author: "Katelyn Vidal"
format: html
---

### Problem 1. Multiple linear regression: model selection and construction (52 points)

Use the information from the `homework-starter-doc.qmd` to do this problem.

# Set up

## reading in packages

```{r}
#| message: false

# general use
library(tidyverse)
library(readxl)
library(here)
library(janitor)

# visualizing pairs
library(GGally)

# model selection
library(MuMIn)

# model predictions
library(ggeffects)

# model tables
library(gtsummary)
library(flextable)
library(modelsummary)

drought_exp <- read_xlsx(path = here("data", 
                                     "Valliere_etal_EcoApps_Data.xlsx"),
                         sheet = "First Harvest")

# quick look at data 
str(drought_exp)
class(drought_exp)
```

## cleaning

```{r}
# cleaning
drought_exp_clean <- drought_exp %>% 
  clean_names() %>% # nicer column names
  mutate(species_name = case_when( # adding column with species scientific names
    species == "ENCCAL" ~ "Encelia californica", # bush sunflower
    species == "ESCCAL" ~ "Eschscholzia californica", # California poppy
    species == "PENCEN" ~ "Penstemon centranthifolius", # Scarlet bugler
    species == "GRICAM" ~ "Grindelia camporum", # great valley gumweed
    species == "SALLEU" ~ "Salvia leucophylla", # Purple sage
    species == "STIPUL" ~ "Nasella pulchra", # Purple needlegrass
    species == "LOTSCO" ~ "Acmispon glaber" # deerweed
  )) %>% 
  relocate(species_name, .after = species) %>% # moving species_name column after species
  mutate(water_treatment = case_when( # adding column with full treatment names
    water == "WW" ~ "Well watered",
    water == "DS" ~ "Drought stressed"
  )) %>% 
  relocate(water_treatment, .after = water) # moving water_treatment column after water
```

# Visualizing

## correlations 

```{r}
ggpairs(drought_exp_clean, # data frame
        columns = c("leaf_dry_weight_g", # columns to visualize
                    "sla", 
                    "shoot_g", 
                    "root_g", 
                    "total_g"), 
        upper = list(method = "pearson")) + # calculating Pearson correlation coefficient
  theme_bw() + # cleaner theme
  theme(panel.grid = element_blank()) # getting rid of gridlines
```

#### a. Make a table *or* list of all the models from class and the last one you constructed on your own. Write a caption for your table. (8 points)

**Caption:** table captions typically go above the table. Number the table and provide a title. Describe what is in the table (columns and rows).

**Table:** In your table, each row should be a model with the model number (1, 2, 3, etc.) and the predictors for each model.

::: callout-note
#### Double check your work!

There should be 5 models total (null model + 4 models with predictors).
:::

#### b. Write a 5-6 sentence "statistical methods" section. (8 points)

Your answer should be in paragraph form and include:

-   how you addressed the central question(s) (i.e. to examine the influence of \_\_\_\_, \_\_\_\_, and \_\_\_\_ on \_\_\_\_\_, I...)\
-   how you chose the final model (i.e. to determine the model that best described \_\_\_\_, I...)\
-   how you visually determined that your final model conformed to the assumptions of a linear model (i.e. to evaluate linear model assumptions, I...)

#### c. Make a visualization of the model predictions with underlying data for your "best" model. (20 points)

Show and annotate all your code. For full credit:

-   make the underlying data more transparent than the model predictions\
-   display species names in full (not as species codes like ENCCAL or ESCCAL)\
-   display water treatment types in full (not as WW or DS)\
-   represent well-watered and drought stressed treatments with different colors\
-   use colors that are *not* the default `ggplot()` colors\
-   facet your plot by species\
-   remove the legend\
-   finalize the plot

::: callout-note
Make sure that the only output is the visualization!
:::

#### d. Write a caption for your visualization. (6 points)

Include a data citation.

#### e. Write a 3-4 sentence results section. (10 points)

Your answer should be in paragraph form and address the following points:

-   what predictors "best" described total mass (include model statistics here)?\
-   on average, what differences did you find between water treatments?\
-   on average, what differences did you find between species?

### Problem 2. Affective visualization (24 points)

In this problem, you will create an **affective visualization** using your personal data in preparation for workshop during week 10.

In lecture, we talked about the three vertices of data visualization: 1) exploratory, 2) affective, and 3) communicative. We've done a lot of exploratory and communicative visualization, but have yet to think about affective visualization.

When thinking of affective visualization, you can expand your ideas of what data visualization *could* be. Some examples of affective visualizations include:

-   [Jill Pelto's paintings](https://www.jillpelto.com/gallery-1)\
-   [Lorraine Woodruff-Long's warming strips quilt](https://www.instagram.com/p/C689Kybxget/?utm_source=ig_web_copy_link&igsh=MzRlODBiNWFlZA==)\
-   [Stefanie Posavec and Giorgia Lupi's Dear Data project](https://www.dear-data.com/all)

**Before starting, update your spreadsheet of observations.**

#### a. Describe in words what an affective visualization could look like for your personal data (3-5 sentences). (2 points)

#### b. Create a sketch (on paper) of your idea. (2 points)

Include a photo of this sketch in your submission.

#### c. Make a draft of your visualization. (12 points)

Feel free to be creative with this! *You do not have to do this in R.* You could create a sculpture, painting, textile object, etc.

If you are making your visualization in R, show the annotated code and the output.

If you are making your visualization outside of R, include a photo of your visualization in your submission.

#### d. Write an artist statement. (8 points)

An artist statement gives the audience context to understand your work. Write 4-5 sentences to address:

-   the content of your piece (what are you showing?)\
-   the influences (what did techniques/artists/etc. did you find influential in creating your work?)\
-   the form of your work (written code, watercolor, oil painting, etc.)\
-   your process (how did you create your work?)

### Problem 3. Statistical critique (36 points)

At this point, you have seen and created a lot of figures for this class. Revisit the paper you chose for your critique and your homework 2, where you described figures or tables in the text. Address the following in full sentences (3-4 sentences each).

**For this section of your homework, you will be evaluated on the logic, conciseness, and nuance of your critique.**

#### a. Revisit and summarize (6 points)

What are the statistical tests the authors are using to address their main research question?

Insert the figure or table you described in Homework 2 here.

#### b. Visual clarity (10 points)

How clearly did the authors *visually* represent their statistics in figures? For example, are the x- and y-axes in a logical position? Do they show summary statistics (means and SE, for example) and/or model predictions, and if so, do they show the underlying data?

#### c. Aesthetic clarity (10 points)

How well did the authors handle "visual clutter"? How would you describe the the data:ink ratio?

#### d. Recommendations (can be longer than 4 sentences, 10 points)

What recommendations would you make to make the figure better? What would you take out, add, or change? Provide explanations/justifications for each of your recommendations.

**Alternatively**, if they did not represent their statistics in a figure, what kind of figure would you recommend to them? Describe the x- and y-axes, with any geometries and aesthetics (colors, etc.). Provide enough explanation such that someone would be able to create the figure you describe in code. Provide explanations/justifications for the figure you would recommend.
